Finch: Sparse and Structured Array Programming with Control Flow

### 2024-04-29 sections/eval/api.tex

Across all three of these kernels, we see that the high level interface for Finch provides a major improvement over DuckDB, ranging from $1.2x-28x$. For triangle counting and SDDMM, this improvement stems from Finch's efficient intersection of the nonzero indices as opposed to DuckDB's use of binary join plans\footnote{This matches with findings in the database literature showing that worst-case optimal joins (which are very similar to our kernel execution) are more efficient than binary joins for these queries \cite{wang2023free}.}. For element-wise operations, this improvement stems from Finch's better handling of expressions which combine index intersection and union and its compressed data representation.

>>>>

Across all three of these kernels, we see that the high level interface for Finch provides a major improvement over DuckDB, ranging from $1.2x-28x$. For triangle counting and SDDMM, this improvement stems from Finch's efficient intersection of the nonzero indices as opposed to DuckDB's use of binary join plans\footnote{This matches with findings in the database literature showing that worst-case optimal joins (which are very similar to our kernel execution) are more efficient than binary joins for these queries \cite{wang_free_2023}.}. For element-wise operations, this improvement stems from Finch's better handling of expressions which combine index intersection and union and its compressed data representation.

### 2024-04-29 sections/eval/spgemm.tex

%Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. 
We compute the $M \times N$ sparse matrix $C$ as the product of $M \times K$ and $K \times N$ sparse matrices $A$ and $B$.
%
There are three main approaches to SpGEMM \cite[Section 2.2]{zhang2021gamma}.
%
The inner-products algorithm takes dot products of corresponding rows and columns, while the outer-products algorithm sums the outer products of corresponding columns and rows.
%
Gustavson's algorithm sums the rows of $B$ scaled by the corresponding nonzero columns in each row of $A$.
%
Inner-products is known to be asymptotically less efficient than the others, as we must do a merge operation to compute each of the $O(MN)$ entries in the output \cite{ahrens2022autoscheduling}.
%
We will show that our ability to implement these latter methods exceeds that of TACO, translating to asymptotic benefits. 
%

>>>>

%Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. 
We compute the $M \times N$ sparse matrix $C$ as the product of $M \times K$ and $K \times N$ sparse matrices $A$ and $B$.
%
There are three main approaches to SpGEMM \cite[Section 2.2]{zhang_gamma_2021}.
%
The inner-products algorithm takes dot products of corresponding rows and columns, while the outer-products algorithm sums the outer products of corresponding columns and rows.
%
Gustavson's algorithm sums the rows of $B$ scaled by the corresponding nonzero columns in each row of $A$.
%
Inner-products is known to be asymptotically less efficient than the others, as we must do a merge operation to compute each of the $O(MN)$ entries in the output \cite{ahrens_autoscheduling_2022}.
%
We will show that our ability to implement these latter methods exceeds that of TACO, translating to asymptotic benefits. 
%
